---
title: "Usage of SQL"
output: html_document
---

# Filtering Rows and Comlums 

With SQL we can filter coloumns with SELECT and rows with WHERE. To show this I will use the Lahman package on RStudio, which is a baseball database. First, we need to need to install the two packages: Lahman and sqldf. 


```{r warning=FALSE, message=FALSE}
library(Lahman)
library(sqldf)
```

Suppose we want to see the total number of homeruns for the Yankees in 1927. We could write the following:

```{r}
query<-"SELECT playerID,yearID,teamID,HR From Batting
Where teamID='NYA' and yearID=1927"
sqldf(query)
```

Batting
colnames(Batting)

Here we will learn how to create a query for New York Yankee players that have more than 40 homeruns.

```{r}
query<-"SELECT playerID,yearID,teamID,HR From Batting
Where teamID='NYA' and HR>=40"
sqldf(query)
```

We will narrow the output more with a query example of players having more than 40 homeruns (HR) but less than 60 strikeouts(SO)

```{r}
query1<-"SELECT playerID,yearID,teamID,SO,HR From Batting
Where SO<60 and HR>40"
sqldf(query1)
```

We are producing the instances of Phillies in the 1970's hitting more than 30 homeruns

```{r}
query2<-"SELECT playerID,yearID,teamID,HR From Batting
Where yearID>1970 and yearID<1979 and HR>=30 and teamID='PHI'"
sqldf(query2)
```

# Ranking Data 

For this example we will present
Player hits with more than 50 homeruns but lets have players with the most homeruns at the top 

```{r}
query3<-"SELECT playerID,yearID,teamID,HR From Batting
Where HR>=50
Order By HR DESC"
sqldf(query3)
```

Find ll instances ofa player striking out less than 10 times. At least 400 at-bat(AB). Players with least SO at top

```{r}
query4<-"SELECT playerID,yearID,teamID,SO, AB From Batting
Where AB>=400 and SO<10
Order by SO"
sqldf(query4)
```

# Aggregation

Aggreation is the clustering of pieces to output one final result. This can be achieved my using grouping and Sum and average. Continuing with our use of the Lahman database we will:

Find Babe Ruth's career homerun total and squash it into one
no comma after homerun
```{r}
query5<-"select playerID, sum(HR) From Batting
WHERE playerID='ruthba01'
Group by playerID"

sqldf(query5)
```

Find the homerun totals of all players but limit the display to only those that hit 600 or 
more. The players with the highest total at the top
where is filtering before group and having is grouping after
```{r}
query6<-"select playerID, sum(HR) From Batting
Group by playerID
Having sum(HR)>600
Order by sum(HR) Desc"

sqldf(query6)
```

# Ordering

Here we will find the players with the highest homerun average over their career.
We will filter out those who have an average of more than 30 HR showing players with the best 
average at the top
avg() #max() #min()
```{r}
query7<-"select playerID, avg(HR) From Batting
Group by playerID
Having avg(HR)>30
Order by avg(HR) Desc"

sqldf(query7)
```
# Joins
This is the joining of two tables using the common variable in both tables using Inner Join and "On Table1.variable=Table2.variable."  When mentioning the variable I have to specify which table it is coming from as seen in the Where table.variable.

```{r}
query8<-"select nameFirst, nameLast, yearID, teamID, HR 
From Batting Inner Join Master
On Batting.playerID=Master.playerID
Where Batting.playerID='ruthba01'"
sqldf(query8)
```

Find all instances where players hitting more than 50 homeruns give their full name
```{r}
query9<-"Select nameFirst, nameLast, yearID, teamID, HR 
From Batting Inner Join Master
On Batting.playerID=Master.playerID
Where Hr>50
order by HR Desc"
sqldf(query9)
```

Our next example is showing Babe Ruth with his team name using the Teams database

```{r}
query10<-"Select playerID, Batting.yearID, name, Batting.HR 
From Batting Inner Join Teams 
On Batting.teamID=Teams.teamID AND Batting.yearID=Teams.yearID
Where playerID='ruthba01'
order by Batting.HR Desc"
sqldf(query10)
```

For our next example we will list all the players named Bob who averaged more than one million dollars per year in salary for their career, including Last Name. 
We will use colnames: Master and Salaries

```{r}
query11<- "Select nameFirst,nameLast,nameLast,avg(salary) 
From Master Inner Join Salaries
On Master.playerID=Salaries.playerID
Where nameFirst='Bob'
Group by Master.playerID
Having avg(salary)>1000000"
sqldf(query11)
```