---
title: "SQL"
output: html_document
---

##filtering rows and comlums 
with Sql you can filter coloumns with Select and Rows with Where. Let us look at an example wusing Lahman baseball database. First, we need to need to load some packages. 

```{r warning=FALSE, message=FALSE}
library(Lahman)
library(sqldf)
```
r
Suppose we want to see the total number of homeruns for the Yankees in 1927. We could write the following
```{r}
query<-"SELECT playerID,yearID,teamID,HR From Batting
Where teamID='NYA' and yearID=1927"
sqldf(query)
```
Batting
colnames(Batting)

Here we will learn how to create a query for New York Yankee player that have more than 40 homeruns
```{r}
query<-"SELECT playerID,yearID,teamID,HR From Batting
Where teamID='NYA' and HR>=40"
sqldf(query)
```
Examples of player having more than 40HR but less than 60strikeouts(SO)
```{r}
query1<-"SELECT playerID,yearID,teamID,SO,HR From Batting
Where SO<60 and HR>40"
sqldf(query1)
```
Instances of Phillies in the 1970's hitting more than 30 homeruns
```{r}
query2<-"SELECT playerID,yearID,teamID,HR From Batting
Where yearID>1970 and yearID<1979 and HR>=30 and teamID='PHI'"
sqldf(query2)
```
Player hits more than 50 homeruns but lets have players with the most homeruns at the top 
```{r}
query3<-"SELECT playerID,yearID,teamID,HR From Batting
Where HR>=50
Order By HR DESC"
sqldf(query3)
```
Find ll instances ofa player striking out less than 10 times. At least 400 at-bat(AB). Players with least SO at top
```{r}
query4<-"SELECT playerID,yearID,teamID,SO, AB From Batting
Where AB>=400 and SO<10
Order by SO"
sqldf(query4)
```